Geddy Lucier
03_Thesis Extensions

This document is designed to test extended hypothesis for the Natural Gas Algonquin Hypothesis

Data Load-in preamble
```{r}
library(tidyverse)
library(dplyr) # Data Manipulation
library(lubridate) # Date manipulation
library(ggplot2) # Visualization
library(riem) # Weather api from airport observations
library(stargazer)


setwd("~/Documents/GitHub/Algonquin_NatGas/data/")

prices <- read_csv("prices_new.csv") # comes from Bloomberg Terminal. 
electricity_demand <- read_csv("electricity_demand.csv")
mix_load_df <- read_csv("mix_load_df.csv")
```
Initial Price Cleaning
```{r}
chicago <- prices[c(1,2)] %>% rename("Date" = "Date...1")
algonquin <- prices[c(3,4)]  %>% rename("Date" = "Date...3")

master_dates <- chicago %>%
  select("Date") %>%
  distinct()

prices <- chicago %>%
  left_join(algonquin, by = "Date")

start_date <- tail(chicago$Date, 1)
end_date <- head(chicago$Date,1)
```
Hypothesis:

The amount of renewable energy generation across the board is correlated with whether or not Algonquin trades or not. Test using a logistic/logit regression



```{r}
boston_mix <- mix_load_df %>%
  filter(respondent == "ISNE" & class == "green") %>%
  select(date, total_renewable, pct_of_total)
  
  
boston_df <- prices %>%
  select(Date,boston) %>% mutate(
  trade_day = ifelse(is.na(boston), 0, 1) 
  #1. Column traded (built off of is.na(algonquin$boston and if there's a price))
  ) %>%
  left_join(boston_mix, by = c("Date" = "date")) %>%
  #2. merge with electricity generation data 
  
  select(trade_day, total_renewable, pct_of_total)
#3. logit regression
model <- glm(trade_day ~ total_renewable + pct_of_total, data = boston_df, family = binomial())

summary(model) 
plot(model)

```
Convert LogOdds to probabilities
```{r}
model$coefficients

b2 <- -6.184141e-05
b3 <-  1.278046e-01 

plogis(b2) # 0.4999845
plogis(b3) # 0.5319077

# Neither of these are particularly predictive in whether there will be trading on a certain day. IE, whether there is demand for natgas at algonquin or not
```
```{r}
ggplot(boston_df, aes(x = pct_of_total, y = trade_day)) +
  geom_point(alpha = 0.5) +  # Add points with some transparency to handle overplotting
  geom_smooth(method = "glm", method.args = list(family = binomial), se = FALSE) +
  labs(title = "Logistic Regression of Trade Day on Total Renewable",
       x = "Total Renewable Energy",
       y = "Trade Day (0 or 1)") +
  theme_minimal()
```
