Geddy Lucier
ENVS - 

Testing out what happens to stock market returns after major market crashes using an event series method.
Using quantmod package for stock data. This is how I learned a lot of my machine learning techniques, so I think it will be great practice

Did S&P 500 returns change significantly after March 16, 2020 (COVID crash) compared to before?

Preamble
```{r}
library(quantmod)
library(ggplot2)
library(dplyr)
library(lubridate)
```
Data preparation



Data loading
```{r}
getSymbols("^GSPC", from = "2019-01-01", to = "2020-09-01")
sp500 <- data.frame(date = index(GSPC), coredata(GSPC))
sp500 <- sp500 %>%
  mutate(return = 100 * (GSPC.Adjusted / lag(GSPC.Adjusted) - 1))

#comes fromr the quantmod package, pulling s&p500 daily prices on the designated dates
```


Event series
```{r}
# building a date to run event series on. Covid crash
event_date <- as.Date("2020-03-16")

pre_window <- sp500 %>% filter(date >= event_date - 60 & date < event_date)
post_window <- sp500 %>% filter(date > event_date & date <= event_date + 60)

```

```{r}
ggplot(sp500 %>% filter(date >= event_date - 60 & date <= event_date + 60),
       aes(x = date, y = return)) +
  geom_line(color = "steelblue") +
  geom_vline(xintercept = as.numeric(event_date), color = "red", linetype = "dashed") +
  labs(title = "S&P 500 Returns Around COVID-19 Crash", y = "Daily Return (%)")
#visualisation
```

```{r}
t_test_result <- t.test(post_window$return, pre_window$return, alternative = "two.sided")
t_test_result
#not a statistically significant result/event
#how does causality/correlation work in this case
```
```{r}
avg_returns <- data.frame(
  period = c("Pre-event", "Post-event"),
  mean_return = c(mean(pre_window$return, na.rm = TRUE),
                  mean(post_window$return, na.rm = TRUE))
)

ggplot(avg_returns, aes(x = period, y = mean_return, fill = period)) +
  geom_col(width = 0.6) +
  labs(title = "Mean Returns Before and After COVID Crash", y = "Mean Daily Return (%)") +
  theme_minimal()
# this seems like a question of causality and market growth rather than anything caused primarily by covid.
#doesn't quite pass the smell test 
#for hurricanes/oil spread, I would need to be very careful about what I 


```





