Geddy Lucier
ENVS - 

Testing out what happens to stock market returns after major market crashes using an event series method.
Using quantmod package for stock data. This is how I learned a lot of my machine learning techniques, so I think it will be great practice for actually applying it to my dataset. 


For this project, I downloaded historical S&P 500 stock market data using the quantmod package, which retrieves financial data directly from Yahoo Finance.
The dataset spans from January 2019 to September 2020 — a period that includes the COVID-19 market crash of March 2020.
This allows me to test whether stock returns changed significantly after the crash. Did S&P 500 returns change significantly after March 16, 2020 (COVID crash) compared to before?

Preamble
```{r}
library(quantmod)
library(ggplot2)
library(dplyr)
library(lubridate)
```
The S&P 500 dataset includes daily financial information for the U.S. stock market index.
Each observation represents one trading day.
Key variables include:

date: the trading day

GSPC.Open, GSPC.High, GSPC.Low, GSPC.Close, GSPC.Volume, GSPC.Adjusted: standard OHLC and volume data

return: calculated daily percentage change in adjusted closing price

The scope covers a 21-month window surrounding the onset of the COVID-19 pandemic, capturing both pre-crash and post-crash behavior.

Data loading
```{r}
getSymbols("^GSPC", from = "2019-01-01", to = "2020-09-01")
sp500 <- data.frame(date = index(GSPC), coredata(GSPC))
sp500 <- sp500 %>%
  mutate(return = 100 * (GSPC.Adjusted / lag(GSPC.Adjusted) - 1))

#comes fromr the quantmod package, pulling s&p500 daily prices on the designated dates
```
I conducted an event study using a two-sample t-test to evaluate whether mean daily returns changed significantly after March 16, 2020 (the height of the COVID-19 market crash) compared to the 60 trading days before. I expect there won't be anything. 

Event study
```{r}
# building a date to run event series on. Covid crash
event_date <- as.Date("2020-03-16")

pre_window <- sp500 %>% filter(date >= event_date - 60 & date < event_date)
post_window <- sp500 %>% filter(date > event_date & date <= event_date + 60)

```

```{r}
ggplot(sp500 %>% filter(date >= event_date - 60 & date <= event_date + 60),
       aes(x = date, y = return)) +
  geom_line(color = "steelblue") +
  geom_vline(xintercept = as.numeric(event_date), color = "red", linetype = "dashed") +
  labs(title = "S&P 500 Returns Around COVID-19 Crash", y = "Daily Return (%)")
#visualisation
```
The two-sample t-test is appropriate because it compares mean values between two independent groups (returns before vs. after the event).
The goal is to determine if the event (COVID crash) led to a statistically significant shift in daily stock returns.
This test assumes approximate normality and independence between daily observations, which is reasonable for short-term return data.

```{r}
t_test_result <- t.test(post_window$return, pre_window$return, alternative = "two.sided")
t_test_result
#not a statistically significant result/event
#how does causality/correlation work in this case
```
The t-test produced a p-value greater than 0.05, indicating that the difference in average daily returns before and after March 16, 2020 was not statistically significant at the 5% level. It was around 0.19. 

While visually the crash appears dramatic, the test suggests that the market’s average daily returns in the 60-day windows were not statistically distinguishable, potentially because the recovery in April–May offset the extreme losses in March.
This highlights that a single-event comparison may not fully capture the volatility or structural changes caused by the pandemic, and so I'll have to see how it handles my data and my dataset. 


```{r}
avg_returns <- data.frame(
  period = c("Pre-event", "Post-event"),
  mean_return = c(mean(pre_window$return, na.rm = TRUE),
                  mean(post_window$return, na.rm = TRUE))
)

ggplot(avg_returns, aes(x = period, y = mean_return, fill = period)) +
  geom_col(width = 0.6) +
  labs(title = "Mean Returns Before and After COVID Crash", y = "Mean Daily Return (%)") +
  theme_minimal()
# this seems like a question of causality and market growth rather than anything caused primarily by covid.
#doesn't quite pass the smell test 
#for hurricanes/oil spread, I would need to be very careful about what I 


```





